---
title: "Analysis summary"
author: "Gordana"
date: '`r Sys.Date()`'
output:
  html_document: default
  word_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(rpart.plot)
library(grpreg)
library(glmnet)
library(table1)
library(OptimalCutpoints)
library(kableExtra)
```



Load fitted models

```{r}
# load fitted models
load( file ="../results/diab_tree")
load( file ="../results/diab_lasso")
load( file ="../results/diab_elastic")

```




### Table 1 (see also Table1.csv)
```{r}

diab_full <- read.csv("../data/clean_modsel.csv", stringsAsFactors =TRUE) %>% 
  mutate(charlson = exp(lcharlson) - 1) %>% 
  dplyr::select( -patient_mrn, -ID, -admit_dttm, , - represent) %>% 
  relocate(readmit) %>% # put readmit in first column
  mutate(age = as.integer(age),
         bgl_min=relevel(bgl_min, ">=4"),
         med_surg = factor(med_surg, levels = 1:2, labels = c("Medical", "Surgical")),
         emergency = factor(emergency, levels = 1:2, labels = c("Emerg", "NotEmerg")),
         charlson = as.integer(charlson),
         disadvantage = as.integer(disadvantage),
         age_s = (age - 70)/10, # scaling
         disadvantage_s = (disadvantage-1000)/50,
         readmit = ifelse(readmit==1,"Yes","No")) # scaling

T1 <- table1( ~   age + sex + ethnicity + dmtypes + charlson +
                             insulin + last90days + med_surg+
                             emergency + MaritalStatus + bgl_max + bgl_min + interaction(bgl_max,bgl_min)+
                             disadvantage +  length_of_stay_days | osp_full_name +readmit  ,
              overall = FALSE,
              data=diab_full,
              render.continuous=c( .="Median [Q1, Q3]"))


T1

write.csv(as.data.frame(T1), file = "../results/Table1.csv", row.names = FALSE)

```



### Figure 2
```{r, fig.width=10}
train_per <- read.csv("../results/model_pred_train.csv") %>% 
  mutate(data = "Training - POW 80%") %>% 
  dplyr::select(AUC:UCL, model, data)
test_per <- read.csv("../results/model_pred_test.csv")  %>% 
  mutate(data = "Test - POW 20%")%>% 
  dplyr::select(AUC:UCL, model, data)
validation_per <- read.csv("../results/model_pred_validation.csv")  %>% 
  mutate(data = "Validation - St George")%>% 
  dplyr::select(AUC:UCL, model, data)


all_AUC <- bind_rows(train_per, test_per, validation_per)

Table2 <- all_AUC %>% 
  transmute(data,
            model,
            `AUC (95% CI)` = paste0(round(AUC,3)," (",round(LCL,3),", ", round(UCL,3),")")) 

Table2 




write.csv(Table2, file = "../results/Table2.csv", row.names = F)

```


```{r}
library(forestplot)

models = data.frame(model = c("tree", "gp_lasso","elastic","nnet"),
           model_nice = c("Classification tree", "Group Lasso","Elastic net","Neural network"))


all_AUC <- all_AUC %>% 
  mutate(interval = paste0(format(round(AUC, 3), nsmall = 3), " ",
                          "(",format(round(LCL, 3), nsmall = 3), " - ", 
                          format(round(UCL,3), nsmall = 3),")")) %>% 
  mutate(data_single = ifelse(model == "tree", data, NA)) %>% 
  left_join(models)

png("../results/AUCplot.png", width=8, height=4,units = "in",
    res = 300)


all_AUC |>
  forestplot(labeltext = c(data_single, model_nice, interval),
             mean = AUC,
             lower = LCL,
             upper = UCL,
             zero = NA,
             mar = unit(rep(5, times = 4), "mm"),
             xticks = seq(0.55,0.8, by = 0.05),
             boxsize = 0.2,
            lineheight =  unit(5,"mm"),
             xlab = "AUC (95% CI)",
             lwd.ci	= 2) %>% 
  fp_add_header(data_single = c("Data")|> fp_txt_plain(),
                model_nice = c("Model")|> fp_txt_plain(),
                interval = c("AUC (95% CI)")|> fp_txt_plain()) %>% 
  fp_set_style(align = "lllr",
               hrz_lines = "#999999",
               txt_gp = fpTxtGp(cex=0.85,
                                ticks=gpar(cex=0.8),
                                xlab = gpar(cex = 0.8)),
               axes = gpar(cex = 0.85)) 



dev.off()

```

### Table 3 and 4

 
```{r include = F}
spec_at_sens <- data.frame()

for(model in c("lasso", "elastic", "nnet"))
{
  for(dat in c("train", "test", "validation")){
    spec_at_sens_loc = read.csv(paste0("../results/auc_tab_",model,"_",dat, ".csv")) %>% 
      dplyr::select(-X) %>% 
      mutate(model = model,
            data = dat, selected = 0)
    sel_ind <- c(max(which(spec_at_sens_loc$sensitivities >= 0.25)),
              max(which(spec_at_sens_loc$sensitivities >= 0.50)),
              max(which(spec_at_sens_loc$sensitivities >= 0.75)))
    spec_at_sens_loc$selected[sel_ind] = 1
    
    spec_at_sens <- spec_at_sens %>% 
      rbind(spec_at_sens_loc)
  }
}


tree_spec <- data.frame()
for(dat in c("train", "test", "validation")){
    tree_spec = read.csv(paste0("../results/auc_tab_tree_",dat, ".csv")) %>% 
      dplyr::select(-X) %>% 
      mutate(model = "tree",
            data = dat) %>% 
      rbind(tree_spec) 
}

tree_spec <- tree_spec %>% 
      mutate(selected = ifelse(is.finite(score), 1, 0))

all_roc <- spec_at_sens %>%
  rbind(tree_spec) %>% 
  arrange(data, sensitivities) %>% 
  mutate(data = case_when(data == "train" ~ "Training - POW 80%",
                          data == "test" ~ "Test - POW 20%",
                          data == "validation" ~ "External validation - St George"))






```



```{r}
Table3 <- all_roc %>% 
  filter(selected == 1) %>% 
  dplyr::select(- selected) %>% 
  arrange(data, model) %>% 
  mutate_if(is.numeric, ~ round(., digits = 3))




Table3 <- Table3 %>% 
  arrange(desc(data)) 


Table3%>% 
  filter(data != "Training - POW 80%") %>% 
  mutate(data = substr(data, 1,10))
write.csv(Table3, file = "../results/Table3.csv", row.names = F)


  

```

#### Proportion of patients that need to be seen 


Calulation

- take all POW data (combined) and All St George data
- take score cutoffs from Table 3
- calculate what proportion of patients are greater than the cut-off from table 3, i.e. would be seen. 


```{r}
data_with_pred_train <- read.csv(file = paste0("../results/data_with_pred_train.csv" ))
data_with_pred_test <- read.csv(file = paste0("../results/data_with_pred_test.csv" ))
data_with_pred_STG <- read.csv(file = paste0("../results/data_with_pred_validation.csv" ))

data_with_pred_POW = bind_rows(data_with_pred_train, data_with_pred_test)


training_scores <- Table3 %>% 
  filter(data == "Training - POW 80%")

training_scores$percent_patients = NA

training_scores$percent_patients[training_scores$model== "elastic"]<- with(training_scores %>% filter(model == "elastic"),
     apply(as.matrix(score), 1, function(x) mean(data_with_pred_POW$pred_elastic > x)))

training_scores$percent_patients[training_scores$model== "lasso"]<- with(training_scores %>% filter(model == "lasso"),
     apply(as.matrix(score), 1, function(x) mean(data_with_pred_POW$pred_lasso > x)))

training_scores$percent_patients[training_scores$model== "nnet"]<- with(training_scores %>% filter(model == "nnet"),
     apply(as.matrix(score), 1, function(x) mean(data_with_pred_POW$pred_nnet > x)))

training_scores$percent_patients[training_scores$model== "tree"]<- with(training_scores %>% filter(model == "tree"),
     apply(as.matrix(score), 1, function(x) mean(data_with_pred_POW$pred_tree > x)))



validation_scores <- Table3 %>% 
  filter(data == "External validation - St George")
validation_scores$score = training_scores$score


validation_scores$percent_patients[validation_scores$model== "elastic"]<- with(validation_scores %>% filter(model == "elastic"),
     apply(as.matrix(score), 1, function(x) mean(data_with_pred_STG$pred_elastic > x)))

validation_scores$percent_patients[validation_scores$model== "lasso"]<- with(validation_scores %>% filter(model == "lasso"),
     apply(as.matrix(score), 1, function(x) mean(data_with_pred_STG$pred_lasso > x)))

validation_scores$percent_patients[validation_scores$model== "nnet"]<- with(validation_scores %>% filter(model == "nnet"),
     apply(as.matrix(score), 1, function(x) mean(data_with_pred_STG$pred_nnet > x)))

validation_scores$percent_patients[validation_scores$model== "tree"]<- with(validation_scores %>% filter(model == "tree"),
     apply(as.matrix(score), 1, function(x) mean(data_with_pred_STG$pred_tree > x)))


prop_screened = bind_rows(training_scores, validation_scores) %>% 
  dplyr::select(data,score,sensitivities,model, percent_patients) %>% 
  mutate(data = ifelse(data == "Training - POW 80%", "POW", "St George"))

prop_screened

write.csv(prop_screened, file = "../results/Table3B.csv", row.names = F)


```




#### Tree

Note - The  circles should have the probability of readmission in that group, however this plot is automatically done from the model, and has weighted sample not the raw data, so it is very overestimated.  The true probabilities are below. There is no way to modify this plot in R, will need to be replotted manually in say Word, with the probabilities swapped out from below. 


```{r}
load( file ="../results/diab_tree")
rpart.plot(diab_tree, type=4, extra=7,box.palette=0)


```

Lookupt table for true for corrected probabilities
```{r}

# training
diab_train <- read.csv("../data/diabetes_clean_train.csv", stringsAsFactors =TRUE) %>%
  mutate(charlson = exp(lcharlson) - 1) %>%
  dplyr::select( -patient_mrn, -ID, -admit_dttm, -osp_full_name, - represent) %>%
  relocate(readmit) %>% # put readmit in first column
  mutate(age = as.integer(age), # for tree model
         bgl_min=relevel(bgl_min, ">=4"),
         med_surg = factor(med_surg, levels = 1:2, labels = c("Medical", "Surgical")),
         emergency = factor(emergency, levels = 1:2, labels = c("Emerg", "NotEmerg")),
         emergency = relevel(emergency, 2),
         charlson = as.integer(charlson), # for tree model
         disadvantage = as.integer(disadvantage), # for tree model
         length_of_stay_days = as.integer(length_of_stay_days),
         ethnicity = relevel(ethnicity, "aust"),
         age_s = (age - 70)/10, # scaling
         age,
         lcharlson_s = 2*(lcharlson - 0.5), # scaling
         disadvantage_s = (disadvantage-1000)/50,
         log_los_days_s = log(length_of_stay_days) - 1.5) # scaling

rel_01= table(diab_train$readmit)
w0=rel_01[2]/rel_01[1]
phat=diab_tree$frame
a=w0*phat$yval2[,5]/(1-phat$yval2[,5])
data.frame(var = phat$var,  uncorrected = round(phat$yval2[,5], 2), corrected = round(a/(1+a),2))

```


### Table 2 - Variable importance

Elastic net and group lasso

```{r}

X <- model.matrix(~ age_s + sex + ethnicity + dmtypes + lcharlson + 
                    insulin + last90days + med_surg+
                    emergency + MaritalStatus + 
                    bgl_max+bgl_min + disadvantage_s + log_los_days_s,
                  data = diab_train)[,-1]



Table4 <- data.frame(gp_lasso_coef = as.matrix(coef(diab_lasso, s = "lambda.min"))[-1, 1] * apply(X, 2, sd),
                     elastic_coef = as.matrix(coef(diab_elastic, s = "lambda.min"))[-1, 1] * apply(X, 2, sd)) %>% 
  transmute(`Group lasso coefficient` = round(gp_lasso_coef,3),
            `Elastic net coefficient` = round(elastic_coef,3)) %>%  
  mutate(`Group lasso coefficient` = ifelse(`Group lasso coefficient` == 0, NA, `Group lasso coefficient`),
         `Elastic net coefficient` = ifelse(`Elastic net coefficient` == 0, NA, `Elastic net coefficient`)) %>% 
  mutate(glasso_rank = rank(-abs(`Group lasso coefficient`)),
         elastic_rank = rank(-abs(`Elastic net coefficient`)))

Table4 
write.csv(Table4, file = "../results/Table4.csv")

```

Neural net



```{r}
Table6 <- read.csv("../results/model/var_imp_results.csv") %>% 
  mutate(nnet_rank = rank(-abs(var_imp_final)))
Table6 %>% 
  arrange(nnet_rank)

write.csv(Table6, file = "../results/Table6.csv", row.names = F)

```



 tree


```{r}

tree_import = t(t(diab_tree$variable.importance)) %>% 
  as.data.frame() %>% 
  mutate(tree_rank = rank(-abs(V1))) %>% 
  dplyr::select(-V1)
tree_import
```

### Misc

#### Flagged patients

Overlap of patients with high/low BGL and those flagged for review by models

```{r}


data_with_pred_POW = data_with_pred_POW %>% 
  mutate(bgl_lt5gt15_both = ifelse(bgl_max == ">15" & bgl_min == "<4","both","no"),
         bgl_lt5gt15_either = ifelse(bgl_max == ">15" | bgl_min == "<4","either","no"),
         osp_full_name = "Prince of Wales Hospital",
         lasso_50 = ifelse(pred_lasso>-2.361,"screen","no"),
         lasso_75 = ifelse(pred_lasso>-2.893,"screen","no"),
         lasso_25 = ifelse(pred_lasso>-1.911,"screen","no"))


with(data_with_pred_POW, table(bgl_lt5gt15_both, lasso_50))
with(data_with_pred_POW, table(bgl_lt5gt15_either, lasso_50))
with(data_with_pred_POW, table(bgl_lt5gt15_both, lasso_75))
with(data_with_pred_POW, table(bgl_lt5gt15_either, lasso_75))
with(data_with_pred_POW, table(bgl_lt5gt15_both, lasso_25))
with(data_with_pred_POW, table(bgl_lt5gt15_either, lasso_25))

with(data_with_pred_POW, prop.table(table(bgl_lt5gt15_either, readmit),margin = 1))
with(data_with_pred_POW, prop.table(table(lasso_50, readmit),margin = 1))

with(data_with_pred_POW, prop.table(table(lasso_25, readmit),margin = 1))
with(data_with_pred_POW, prop.table(table(lasso_75, readmit),margin = 1))

```

#### Raw coeficients for equations

```{r }

Table5 <- data.frame(gp_lasso_coef = as.matrix(coef(diab_lasso, s = "lambda.min"))[, 1] ,
                     elastic_coef = as.matrix(coef(diab_elastic, s = "lambda.min"))[, 1] ) %>% 
  # filter(!(gp_lasso_coef == 0 & elastic_coef == 0))%>%  
  rownames_to_column(var = "predictor") %>% 
  pivot_longer(2:3, names_to = "model") %>% 
  pivot_wider(names_from = predictor, values_from = value) %>% 
  mutate(age = age_s/10,
         disadvantage = disadvantage_s/50,
         log_los_days = log_los_days_s,
         med_surgMed = -med_surgSurgical,
         sex_Male = -sexNot_Male,
         insulinInsulin  = - insulinNo_insulin ,
         Intercept = `(Intercept)`-70*age - 1000*disadvantage - 1.5*log_los_days - med_surgMed - sex_Male - insulinInsulin) %>% 
  relocate(age, .before = age_s) %>% 
  relocate(disadvantage, .before = disadvantage_s) %>% 
  relocate(log_los_days, .before = log_los_days_s) %>% 
  relocate(Intercept, .before = `(Intercept)`) %>% 
  dplyr::select(-age_s, -disadvantage_s, -log_los_days_s, -`(Intercept)`, -med_surgSurgical,
                - sexNot_Male, - insulinNo_insulin) 
  

Table5 <- t(Table5[,-1])

### calculate odds ratios here



colnames(Table5) = c('Group lasso coefficient', 'Elastic net coefficient')
Table5 <- as.data.frame(Table5) %>% 
  mutate(GL_or = exp(`Group lasso coefficient`),
         EN_or = exp(`Elastic net coefficient`))

Table5
write.csv(Table5, file = "../results/Table5.csv")

```




```{r, echo=FALSE, eval = FALSE}

#This checks the above corrections to the formula are correct.

coef_lasso_mod <- as.matrix(coef(diab_lasso, s = "lambda.min"))
X_mod <- model.matrix(~ age_s + sex + ethnicity + dmtypes + lcharlson + insulin + last90days + med_surg + emergency + MaritalStatus + 
                        bgl_max  + bgl_min + disadvantage_s + log_los_days_s,
                  data = diab_train)

cbind(rownames(coef_lasso_mod),colnames(X_mod))

mod_pred = X_mod%*%coef_lasso_mod

diab_train_raw = diab_train %>% 
  mutate(med_surg = relevel(as.factor(med_surg), ref = "Surgical"),
         sex = relevel(as.factor(sex), ref = "Not_Male"),
         insulin = relevel(as.factor(insulin), ref = "No_insulin"))


coef_lasso_raw = as.matrix(Table5[,1])
X_raw <- model.matrix(~  age + ethnicity + dmtypes + lcharlson + last90days + emergency + MaritalStatus + bgl_max + bgl_min + 
                        disadvantage + log(length_of_stay_days) + med_surg + sex + insulin,
                      data = diab_train_raw)
cbind(rownames(coef_lasso_raw),colnames(X_raw))

raw_pred = X_raw%*%coef_lasso_raw

all.equal(mod_pred, raw_pred)

```






#### Readmission rates 

Per patient (based on last admission)

```{r}
diabetes_full <- read.csv("../data/clean_modsel.csv")


diabetes_full %>% 
  group_by(osp_full_name) %>% 
  summarise(patients_readmitted = mean(readmit),
            patients_represented = mean(represent))

```

Per admission

```{r}

diabetes_alladmit <- read.csv("../data/clean_alladmit.csv")


diabetes_alladmit %>% 
  group_by(osp_full_name) %>% 
  summarise(admissions_readmitted = mean(readmit),
            admissions_represented = mean(represent))



```



## References

Agresti, Alan. Categorical data analysis. Vol. 792. John Wiley & Sons, 2012.

Altmann, A., Toloşi, L., Sander, O., & Lengauer, T. (2010). Permutation importance: a corrected feature importance measure. Bioinformatics, 26(10), 1340-1347.

Terry Therneau and Beth Atkinson (2019). rpart: Recursive Partitioning and Regression Trees. R package
  version 4.1-15. https://CRAN.R-project.org/package=rpart
  
Breheny P and Huang J (2015). Group descent algorithms for nonconvex penalized linear and logistic
  regression models with grouped predictors. Statistics and Computing, 25: 173-187.
  
Schisterman, E.F., Perkins, N.J., Liu, A. and Bondell, H. (2005). Optimal cutpoint and its corresponding Youden index to discriminate individuals using pooled blood samples. Epidemiology 16, 73–81.

Jerome Friedman, Trevor Hastie, Robert Tibshirani (2010). Regularization Paths for Generalized Linear
  Models via Coordinate Descent. Journal of Statistical Software, 33(1), 1-22. URL
  http://www.jstatsoft.org/v33/i01/.
  
Monica Lopez-Raton, Maria Xose Rodriguez-Alvarez, Carmen Cadarso Suarez, Francisco Gude
  Sampedro (2014). OptimalCutpoints: An R Package for Selecting Optimal Cutpoints in
  Diagnostic Tests. Journal of Statistical Software, 61(8), 1-36. DOI 10.18637/jss.v061.i08.
  
  Intrator O. and Intrator N. (1993) Using Neural Nets for Interpretation of Nonlinear Models. Proceedings of the Statistical Computing Section, 244-249 San Francisco: American Statistical Society (eds.)